# Reproducible Research: Assignment 1

## 1) Loading and preprocessing the data

``` {r} 
activity<-read.csv("~/Desktop/Data_Science/Coursera_DS/activity.csv")
```

## 2) What is mean total number of steps taken per day?

```{r} 
steps<-aggregate(activity[,1],by=list(activity[,2]),FUN=sum)
# ignoring NA values
final.steps<-steps[complete.cases(steps),]
# histogram
barplot(final.steps$x, 
        names.arg = final.steps$Group.1, 
        main = "Total Steps (Daily)",
        xlab = "Date", 
        ylab = "Steps")
# report mean and median
summary(final.steps$x)
```

## 3) What is the average daily activity pattern?

``` {r}
interval<-aggregate(steps ~ interval, activity, mean)

#plotting
plot(interval, 
     type='l', 
     main="Daily Activity", 
     xlab="Interval", 
     ylab="Average Steps")

#calculating max interval for max Steps
max.steps <- which.max(interval$steps)
interval$interval[max.steps]
```

## 4) Imputing missing values

``` {r}
#NA count
sum(is.na(activity$steps))
```

To fill in the missing values, we will be using the mean per given interval. The new dataset is called 'new.activity'.

``` {r}
require(plyr)
require(Hmisc)
new.activity <- ddply(activity, ~ interval, transform, steps = impute(steps,mean))


#reordering rows in descending order, because ddply() sorts them by interval
new.activity<-new.activity[row.names(new.activity),]

#plotting
new.steps <- aggregate(steps ~ date, new.activity, sum)
barplot(new.steps$steps, names.arg = new.steps$date, main = "New Total Steps (Daily)",xlab = "Date", ylab = "Steps")

#reporting
mean(new.steps$steps)
median(new.steps$steps)
```

The methodology used to fill in the missing values does not impact the results.

## 5) Are there differences in activity patterns between weekdays and weekends?

``` {r}
#creating new factor weekdays / weekend.
new.activity$date <- as.Date(new.activity$date)
new.activity$date <- factor((weekdays(new.activity$date) %in% 
        c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')), 
        levels=c(FALSE, TRUE), labels=c('weekend', 'weekday') )

#plotting
require(dplyr)
new.plot<-new.activity %>% group_by(interval,date) %>% summarise(mean = mean(steps))
require("lattice")
xyplot(mean ~ interval | date,data = new.plot, 
       type = "l", 
       layout = c(1,2), 
       xlab = "Interval", 
       ylab = "Number of steps", 
       main = "")
```
